---
title: "MSDS-6373-Project1"
author: "Ryan H, Jason M"
date: "5/30/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(stringr)
library(plotly)
library(ggplot2)
library(plyr) # Load first before dplyer 
library(dplyr)
library(tidyr)

# Turn off scientific notation. Mainly for the Percentage Expenditure
options(scipen = 999)
```

Read in data first 
```{r}
life_exp_loc <- 'Data/LifeExpectancyData.csv'
# Create the data frame that will not be modified 
life_exp <- read.csv(life_exp_loc, sep=',', header = TRUE) 
```

# EDA
### Start looking at missing data to get an idea of what to expect 
```{r}
# Create data frame of rows that contain missing data 
rows_with_na <- life_exp[rowSums(is.na(life_exp)) > 0,] # 1,289 rows...

# Count how many na's per column
cols_with_na <- life_exp[is.na(life_exp$Schooling),]

# Columns with # of 'na' values
# -----------------------------
# Life.expectancy: 10 *
# Adult.Mortality: 10
# Alcohol: 194
# Hepatitis.B: 553 *
# BMI: 34
# Polio: 19
# Total.expenditure: 226
# Diphtheria: 19
# GDP: 448
# Population: 652 *
# thiness1-19years: 34
# thiness5-9years: 34
# Income.comp.of.resc: 167
# Schooling: 163
# ----------------------------- 
```
- The two biggest column with the most amount of data missing is Hepatitis.B and the 
Population. 
- Since Life.expectancy is the value we are trying to predict, the rows that have 
missing values for Life.expectancy should not be used in the training model
- Will need to do some more EDA and keep an eye on these columns to see how 
big of an impact they have.  
- A lot of 0 values for infant deaths... That doesn't seem right. Looks like something
that could throw off the data. 
- In general keep an eye out if mortality rate in general has any affect
- Also keep an eye on Percentage.expenditure. It's zero for a lot of countries that
are considered developed, and there is another column for total spending on health
for each country. 

### Create a subset of data without the rows containing 'na' in Life.expectancy 
```{r}
# New DF with no NA Life.expectrancies 
life_exp_v002 <- life_exp[!is.na(life_exp$Life.expectancy),]
```

Make some modifications to the columns. Changing Names of columns to be more relative
and change the data types of some columns 
```{r}
# New version
life_exp_v003 <- life_exp_v002

# I know I can write this as one line, but this reads better
colnames(life_exp_v003)[9] <- c("Hep.B.Shots")
colnames(life_exp_v003)[10] <- c("Measle_Cases") # Per 1k
colnames(life_exp_v003)[13] <- c("Polio_Shots") # Percent Amoung 1yr olds
colnames(life_exp_v003)[14] <- c("Total_health_Spending") # Percent of total gov expenditure
colnames(life_exp_v003)[15] <- c("Diphtheria_shots") # Percent among 1yr olds
colnames(life_exp_v003)[16] <- c("HIV_AIDS_birth_deaths") # Per 1k

# Change Country, Year, and Status columns to as.factor
tmp_chng_col <- c("Country", "Year", "Status")
life_exp_v003[tmp_chng_col] <- lapply(life_exp_v003[tmp_chng_col], as.factor)
```

Basic data prepping done. We want to now start to look at the data and check for collinearity,
outliers, or possibly data that could benifit from log transformation or data that may repeat.
```{r}
# Create a V004 
life_exp_v004 <- life_exp_v003

# Removing infant death because I feel the under.five.deaths includes infant deaths
life_exp_v004 <- life_exp_v004[-c(6, )]





```

















