---
title: "MSDS-6373-Project1"
author: "Ryan H, Jason M"
date: "5/30/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(stringr)
library(plotly)
library(ggplot2)
library(plyr) # Load first before dplyer 
library(dplyr)
library(tidyr)
library(corrplot)
# Turn off scientific notation. Mainly for the Percentage Expenditure
options(scipen = 999)
```

Read in data first 
```{r}
life_exp_loc <- 'Data/LifeExpectancyData.csv'
# Create the data frame that will not be modified 
life_exp <- read.csv(life_exp_loc, sep=',', header = TRUE) 
```

# EDA
### Start looking at missing data to get an idea of what to expect 
```{r}
# Create data frame of rows that contain missing data 
rows_with_na <- life_exp[rowSums(is.na(life_exp)) > 0,] # 1,289 rows...

# Count how many na's per column
cols_with_na <- life_exp[is.na(life_exp$Schooling),]

# Columns with # of 'na' values
# -----------------------------
# Life.expectancy: 10 *
# Adult.Mortality: 10
# Alcohol: 194
# Hepatitis.B: 553 *
# BMI: 34
# Polio: 19
# Total.expenditure: 226
# Diphtheria: 19
# GDP: 448
# Population: 652 *
# thiness1-19years: 34
# thiness5-9years: 34
# Income.comp.of.resc: 167
# Schooling: 163
# ----------------------------- 
```
- The two biggest column with the most amount of data missing is Hepatitis.B and the 
Population. 
- Since Life.expectancy is the value we are trying to predict, the rows that have 
missing values for Life.expectancy should not be used in the training model
- Will need to do some more EDA and keep an eye on these columns to see how 
big of an impact they have.  
- A lot of 0 values for infant deaths... That doesn't seem right. Looks like something
that could throw off the data. 
- In general keep an eye out if mortality rate in general has any affect
- Also keep an eye on Percentage.expenditure. It's zero for a lot of countries that
are considered developed, and there is another column for total spending on health
for each country. 

### Create a subset of data without the rows containing 'na' in Life.expectancy
### And also seperate the training from test set 85/15
```{r}
# New DF with no NA Life.expectrancies 
life_exp_v02 <- life_exp[!is.na(life_exp$Life.expectancy),]

# Sperate the data 
set.seed(1536)
split_size <- floor(0.85 * nrow(life_exp_v02))
train_ind <- sample(seq_len(nrow(life_exp_v02)), size = split_size)
life_exp_train_v01 <- life_exp_v02[train_ind,]
life_exp_test <- life_exp_v02[-train_ind,]
```

Make some modifications to the columns. Changing Names of columns to be more relative
and change the data types of some columns 
```{r}
# New version
life_exp_train_v02 <- life_exp_train_v01

# I know I can write this as one line, but this reads better
colnames(life_exp_train_v02)[9] <- c("Hep.B.Shots")
colnames(life_exp_train_v02)[10] <- c("Measle_Cases") # Per 1k
colnames(life_exp_train_v02)[13] <- c("Polio_Shots") # Percent Amoung 1yr olds
colnames(life_exp_train_v02)[14] <- c("Total_health_Spending") # Percent of total gov expenditure
colnames(life_exp_train_v02)[15] <- c("Diphtheria_shots") # Percent among 1yr olds
colnames(life_exp_train_v02)[16] <- c("HIV_AIDS_birth_deaths") # Per 1k

# Change Country, Year, and Status columns to as.factor
tmp_chng_col <- c("Country", "Year", "Status")
life_exp_train_v02[tmp_chng_col] <- lapply(life_exp_train_v02[tmp_chng_col], as.factor)
```

Basic data prepping done. We want to now start to look at the data and check for collinearity,
outliers, or possibly data that could benifit from log transformation or data that may repeat.
```{r, figures-side, fig.show="hold", out.width="32%"}
# Create a v03 
life_exp_train_v03 <- life_exp_train_v02

# Loop through the categories and plot them to get a better understanding of the distribution
col_names <- names(life_exp_train_v03)
col_names_index <- 1

for (col in life_exp_train_v03) {
  plot(life_exp_train_v03$Life.expectancy, col, 
       main=col_names[col_names_index], xlab="Life Expectancy")
  col_names_index <- col_names_index + 1
}
```

### Notes:

- Infant Deaths and under.five.deaths are strongly related
- BMI almost looks like it can be broken down into 3 categories 
- Population has some extreme outliers. Further research says it's all from India
- There is some commonality between charts for categories that record immunization (shots)
- Percentage.expenditure, GDP look to be quadratic 

Apply a log transformation to the 2 charts that look to be quadratic. 
Also drop the 5-9 Thinness. It's the same chart as the 10-19 thinness chart. 
```{r, fig.show="hold", out.width="50%"}
# Version v04
life_exp_train_v04 <- subset(life_exp_train_v03, select=-thinness.5.9.years) # Removes Thiness 5-9 col

# Log transform Percentage.expenditure, GDP,
life_exp_train_v04$percentage.expenditure <- log(life_exp_train_v04$percentage.expenditure)
life_exp_train_v04$GDP <- log(life_exp_train_v04$GDP)

# Check the charts after transformation
# It's only three, I don't need to loop through this 
plot(life_exp_train_v04$Life.expectancy, life_exp_train_v04$percentage.expenditure,
     main="percentage.expenditure", xlab="Life Expectancy")
plot(life_exp_train_v04$Life.expectancy, life_exp_train_v04$GDP,
     main="GDP", xlab="Life Expectancy")

```

Create a correlation table with only the continous data
https://cran.r-project.org/web/packages/olsrr/vignettes/regression_diagnostics.html
```{r}
# Create a version with only numerical values
life_exp_train_v04.1 <- subset(life_exp_train_v04, select=c(-Country, -Year, -Status))

# Using the corrplot package 
# Using pairwise.complete.obs to ignore the na values, but still include the row
cor_num_only <- cor(life_exp_train_v04.1, use='pairwise.complete.obs')

```
There was a strong correlation between

- GDP and Percentage.expenditure (when ran with v004. V005 gave non-real numbers) @.89
- under.five.deaths and infant deaths @.99
- HepB and Diphtheria shots @.61

I believe the path forward will be dropping Percentage.expenditure, infant.deaths because
they have a lower correlation with Life.Expectancy. The shots will stay and the variable 
selection will decide if they stay or go. 

```{r}
# v006
# Remove Percentage.expedeture and infant.deaths
life_exp_train_v05 <- subset(life_exp_train_v04, select=c(-infant.deaths, -percentage.expenditure))
```












